**The statistical analysis of non-linear longitudinal data in biomedical research**

### 2021-01-05 Manuscripts/Manuscript_AM_v1.Rmd
Updated manuscript, html_2 document works now, need to add non-linear anova graph and text and finish summary of GAMs

---
title: '**The statistical analysis of non-linear longitudinal data in biomedical research**'
author: 
- Ariel Mundo^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- Timothy J. Muldoon^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- John R. Tipton^[Department of Mathematical Sciences, University of Arkansas, Fayetteville]

bibliography: refs.bib
output:
  pdf_document: 
    fig_caption: yes
  html_document: default
  css: style.css
  word_document:
    reference_docx: docx_template.docx
mainfont: Times New Roman
link-citations: yes
subtitle: '_Beyond repeated measures ANOVA and Linear Mixed Models_'
csl: ieee.csl
---

## Paper outline
<span style="color: red;">\textcolor{red}{The paper Introduction has been updated, proposed sections appear at the end of the document as well as an initial graph.}</span>

## Background

@@

## Section 1:

### Challenges presented by longitudinal studies

### 1 The "frequentist" case for longitudinal data

@@

#### 1.1 Linear relationship

@@

\begin{equation}
  y = \begin{cases}
  \mu + \beta_1*time_{t}+\epsilon_{ti}   & \mbox{if Group A}\\
  \mu + \beta_1 * time_{t} + \beta_2 * treatment_{j} +\beta_3* time_{t} * treatment_{j}        +\pi_{i}+ \epsilon_{tij}  & \mbox{if Group B}\\
  \end{cases}
  (\#eq:ANOVA-by-group)
\end{equation}

@@

  y = \begin{cases}
  \mu +\boldsymbol{\beta}_1* f(time)  & \mbox{if Group A} \\
  \mu +\boldsymbol{\beta}_1* f(time)  + \beta_2* treatment + \boldsymbol{\beta}_3* f(time) *     treatment   & \mbox{if Group B}\\
  \end{cases}
  \\
  (\#eq:GLM
\end{equation}

@@

```{r,include=FALSE}
## Example with linear response
example <- function(n_time = 6, 
                    fun_type = "linear", 
                    error_type = "correlated") {

@@

```{r, fig.width=16, fig.height=9, out.width="100%", echo=FALSE}
# linear response, correlated errors (subject effect)
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with linear function and correlated errors',
    theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```

#### 1.2 Covariance in rm-ANOVA and LMEMs

@@

#### 1.3 Missing observations

@@

#### 2 GAMs as a special case of Generalized Linear Models

>>>>

---
title: '**The statistical analysis of non-linear longitudinal data in biomedical research**'
author:
- Ariel Mundo^[Department of Biomedical Engineering, University of Arkansas, Fayetteville]
- Timothy J. Muldoon^[Department of Biomedical Engineering, University of Arkansas,
  Fayetteville]
- John R. Tipton^[Department of Mathematical Sciences, University of Arkansas, Fayetteville]
subtitle: _Beyond repeated measures ANOVA and Linear Mixed Models_
bibliography: refs.bib
csl: ieee.csl
css: style.css
output:

  bookdown::html_document2: default
  bookdown::pdf_document2:
    keep_tex: yes
    fig_caption: yes
  bookdown::word_document2: default
  link-citations: true
---

# Background

@@

# Challenges presented by longitudinal studies

## The "frequentist" case for longitudinal data

@@

## Linear relationship

@@

\begin{equation}
  y = \begin{cases}
  \mu + \beta_1*time_{t}+\epsilon_{ti}   & \mbox{if Group A}\\
  \mu + \beta_1 * time_{t} + \beta_2 * treatment_{j} +\beta_3* time_{t} * treatment_{j}+\pi_{i}+\epsilon_{tij}  & \mbox{if Group B}\\
  \end{cases}
  (\#eq:ANOVA-by-group)
\end{equation}

@@

  y = \begin{cases}
  \mu +\boldsymbol{\beta}_1* f(time)  & \mbox{if Group A} \\
  \mu +\boldsymbol{\beta}_1* f(time)  + \beta_2* treatment + \boldsymbol{\beta}_3* f(time) *     treatment   & \mbox{if Group B}\\
  \end{cases}
  (\#eq:GLM)
\end{equation}

@@

```{r,include=FALSE,message=FALSE,echo=FALSE}
## Example with linear response
example <- function(n_time = 6, 
                    fun_type = "linear", 
                    error_type = "correlated") {

@@

```{r, fig.width=16, fig.height=9, out.width="100%", echo=FALSE,message=FALSE}
# linear response, correlated errors (subject effect)
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with linear function and correlated errors',
    theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```
It is clear from this graph that

```{r, fig.width=16, fig.height=9, out.width="100%"}
# quadratic response, correlated errors (subject effect)
plot_example(example(fun_type = "quadratic", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with quadratic function and correlated errors',
    theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```

## Covariance in rm-ANOVA and LMEMs

@@

## Missing observations

@@

# GAMs as a special case of Generalized Linear Models
